---
title: "Subcellular proteomic analysis of the insula in alcohol dependence"
subtitle: "NBIS support {{< 7314 >}}"
author: Yuan Li
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
  pdf_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
editor_options:
  chunk_output_type: inline
params:
  dataset: 1
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)


```

```{r,echo=FALSE,include=FALSE}
# project variables
rep_nbis_id <- "7314"
rep_report_version <- "1.0"
rep_request <- "Philippe Melas"
rep_request_email <- "philippe.melas@ki.se"
rep_pi <- "Philippe Melas"
rep_pi_email <- "philippe.melas@ki.se"
rep_org <- "KI"
rep_nbis <- "Yuan Li"
rep_nbis_email <- "yuan.li@nbis.se"
```

::: boxy
**NBIS ID:** `r rep_nbis_id`\
**Report Version:** `r rep_report_version`\
**Request by:** `r paste0(rep_request," (",rep_request_email,")")`\
**Principal Investigator:** `r paste0(rep_pi," (",rep_pi_email,")")`\
**Organisation:** `r rep_org`\
**NBIS Staff:** `r paste0(rep_nbis," (",rep_nbis_email,")")` **conda env:** TTN
:::

<br>

# Support Request

[7314](https://projects.nbis.se/issues/7314#change-40330) In brief, the data that we have, derive from the following experimental design: Brain samples from alcohol dependent patients (N=12 cases) and matched controls (N=12 controls) were first submitted to subcellular fractionations, which yielded three fractions (i.e., synaptosomes, cytosol, insoluble fraction). All three fractions (total N=72 samples) were subjected to TMT-based quantitative proteomics. Data were provided by a commercial company (Creative Proteomics) in .xlsx format. The company also performed and provided results from basic bioinformatic comparisons (inc. differential expression, pathway analysis, etc.), which revealed some statistical significances between cases and controls on the nominal p-level. We would now like to re-analyze the data and prepare material, e.g., Figures/Tables and Methods, for a publication.

# Data

The data has already been filtered by the company (e.g. removed proteins with CONTA, REV... as prefix of their names, some **potential** contaminating proteins etc. were kept for they may also represent real proteins)

Because there is a big batch effect (between G1 and G2, see the PCA plot in ../results/old/totalAbundanceNormalized/PELLET/differentialAnalysis/Comparisons_loess0.1/Norm-stats-report-Comparisons_loess0.1.pdf), and there are almost no significant protein. So I decided to renormalize the raw intensity, and see how it will affect the results.

**Aim for this particular notebook**: Look for ovelapping genes between our Limma- pldDA-identified differential genes and GWAS studies; look for perturbagens that have reverse expression patterns comparing to the AUD

```{r}
library(readxl)
library(dplyr)
library(metafor)
```

# Import the data and preparation

```{r}
# Import GWAS genes from different studies
library(readxl)
# Zhou et al. 2023: https://www.nature.com/articles/s41591-023-02653-5 # ST1, only European data are used
gwas1_euro = read_excel(
    paste0("../doc/Paper/GWAS/41591_2023_2653_MOESM3_ESM.xlsx"), 
    sheet=3, col_names=TRUE, na="", skip=0)
gwas1_euro
gwas1_cross_ancestry = read_excel(
    paste0("../doc/Paper/GWAS/41591_2023_2653_MOESM3_ESM.xlsx"), 
    sheet=4, col_names=TRUE, na="", skip=0)
gwas1_cross_ancestry
# Saunders et al. 2022: https://doi.org/10.1038/s41586-022-05477-4 # Supplementary Table 1 is used
gwas2 = read_excel(
    paste0("../doc/Paper/GWAS/41586_2022_5477_MOESM3_ESM.xlsx"), 
    sheet=3, col_names=TRUE, na="", skip=0)
gwas2_euro = gwas2 %>% filter(Phenotype=="DrnkWk") %>% filter(Ancestry=="EUR")
gwas2_trans_ancestry = gwas2 %>% filter(Phenotype=="DrnkWk") %>% filter(Ancestry=="Trans-ancestry")
gwas2_euro
gwas2_trans_ancestry
# Icick et al. 2023: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10775504/
gwas3_euro = read_excel(
    paste0("../doc/Paper/GWAS/AUD_GWAS_list_Nat Hum Beh 2025.xlsx"), 
    sheet=2, col_names=TRUE, na="", skip=0)
gwas3_euro
gwas3_multi_ancestry = read_excel(
    paste0("../doc/Paper/GWAS/AUD_GWAS_list_Nat Hum Beh 2025.xlsx"), 
    sheet=3, col_names=TRUE, na="", skip=0)
gwas3_multi_ancestry



# Import Limma differential genes
compartments = c("CYTOSOLIC", "PELLET", "SYNAPTOSOMAL")
limma = lapply(compartments, function(i) {
  dat = read_excel(
    paste0("../results/rerun/DifferentialAnalysis/Differential_Table_",i,".xlsx"), 
                  sheet=1, col_names=TRUE, na="", skip=0)
  dat = dat[, 1:8]
  dat$Fraction = i
  dat$Protein = dat$Protein_IDs
  dat
}) %>% setNames(compartments) 
limma2 = limma %>% Reduce(rbind, .) %>% filter(adj.P.Val<0.2&abs(logFC)>0.22)
limma2 = limma2[!duplicated(limma2$Protein_IDs), ]



# Import DIABLO selected genes:
NAMES = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
loadings = lapply(1:3, function(i) {
  dat = read_excel(
    paste0("../results/rerun/Integration/DIABLO/selectedProteins/SelectedFeatures_weights.xlsx"), 
                  sheet=i, col_names=TRUE, na="", skip=0)
  dat$Fraction = NAMES[i]
  dat
}) %>% Reduce(rbind, .)

```

```{r}
print("Overlapping between Limma differential genes and GWAS study Zhou2023: ")
print("European:")
A1 = intersect(gwas1_euro$nearestGene, limma2$Genes)
A1
# [1] "CADM1" "ACSS3"
print("Cross ancestry: ")
A2 = intersect(gwas1_cross_ancestry$nearestGene, limma2$Genes)
A2
# "ACSS3"

print("Overlapping between Limma differential genes and GWAS study Saunders2022: ")
print("European:")
B1 = intersect(gwas2_euro$Gene, limma2$Genes)
B1
# [1] "CPNE4"   "HNRNPH1" "ACSS3"   "SNRPD3" 
print("Trans_ancestry: ")
B2 = intersect(gwas2_trans_ancestry$Gene, limma2$Genes)
B2
# "CPNE4"   "CTNNA1"  "HNRNPH1" "ACSS3"   "DCLK1"   "SAMHD1" 

print("Overlapping between Limma differential genes and GWAS study Icick2025: ")
print("European:")
E1 = intersect(gwas3_euro$`Credibly mapped gene`, limma2$Genes)
E1
 
print("Multi_ancestry: ")
E2 = intersect(gwas3_multi_ancestry$`Credibly mapped gene`, limma2$Genes)
E2


print("Overlapping between DIABLO selected genes and GWAS study Zhou2023: ")
print("European:")
C1 = intersect(gwas1_euro$nearestGene, loadings$Genes)
C1
# "ACSS3"
print("Cross_ancestry")
C2 = intersect(gwas1_cross_ancestry$nearestGene, loadings$Genes)
C2
# "ACSS3"

print("Overlapping between DIABLO selected genes and GWAS study Saunders2022: ")
print("European:")
D1 = intersect(gwas2_euro$Gene, loadings$Genes)
D1
# "ACSS3"
print("Trans_ancestry")
D2 = intersect(gwas2_trans_ancestry$Gene, loadings$Genes)
D2
# "ACSS3"


print("Overlapping between DIABLO selected genes and GWAS study Icick2025: ")
print("European:")
F1 = intersect(gwas3_euro$`Credibly mapped gene`, loadings$Genes)
F1
print("Multi_ancestry: ")
F2 = intersect(gwas3_multi_ancestry$`Credibly mapped gene`, loadings$Genes)
F2

c(A1, A2, B1, B2, C1, C2, D1, D2, E1, E2, F1, F2) %>% unique() %>% length()

table = data.frame(list(Limma=c(paste(A1,collapse = "/"), paste(A2,collapse="/"), 
                                paste(B1,collapse = "/"), paste(B2,collapse="/"),
                                paste(E1,collapse = "/"), paste(E2,collapse="/")), 
                plsDA=c(paste(C1,collapse="/"), paste(C2,collapse="/"),
                        paste(D1,collapse="/"), paste(D2,collapse="/"),
                        paste(F1,collapse="/"), paste(F2,collapse="/"))))
rownames(table) = c(paste0("Zhou et al. 2023_", c("European", "Multi-ancestry")),
                    paste0("Saunders et a. 2022_", c("European", "Multi-ancestry")), 
                    paste0("Icick et al. 2025_", c("European", "Multi-ancestry")))
table = table %>% tibble::rownames_to_column(var="Study")

dir.create("../results/rerun/Additional/", recursive = T, showWarnings = F)
write.table(table, "../results/rerun/Additional/GWASoverlapping.csv", sep="\t", 
            col.names=T, row.names=F, quote=F)
```

```
[1] "Overlapping between Limma differential genes and GWAS study Zhou2023: "
[1] "European:"
[1] "CADM1" "ACSS3"
[1] "Cross ancestry: "
[1] "ACSS3"
[1] "Overlapping between Limma differential genes and GWAS study Saunders2022: "
[1] "European:"
[1] "CPNE4"   "HNRNPH1" "ACSS3"  
[1] "Trans_ancestry: "
[1] "CPNE4"   "CTNNA1"  "HNRNPH1" "ACSS3"   "DCLK1"   "SAMHD1" 
[1] "Overlapping between Limma differential genes and GWAS study Icick2025: "
[1] "European:"
character(0)
[1] "Multi_ancestry: "
character(0)
[1] "Overlapping between DIABLO selected genes and GWAS study Zhou2023: "
[1] "European:"
[1] "ACSS3"
[1] "Cross_ancestry"
[1] "ACSS3"
[1] "Overlapping between DIABLO selected genes and GWAS study Saunders2022: "
[1] "European:"
[1] "ACSS3"
[1] "Trans_ancestry"
[1] "ACSS3"
[1] "Overlapping between DIABLO selected genes and GWAS study Icick2025: "
[1] "European:"
character(0)
[1] "Multi_ancestry: "
character(0)
[1] 7
```






# Look for perturbagens from L1000 database that show reverse expression pattern comparing to AUD

```{r}
# # Check the data
# GSE70138 = cmapR::parse_gctx("../results/rerun/Additional/DrugTarget_L1000/data/GSE70138_Broad_LINCS_Level5_COMPZ_n118050x12328_2017-03-06.gctx")
# 
# grep("_NEU_", colnames(GSE70138@mat), value=T) %>% sapply(., function(i) {stringr::str_split(i, ":")[[1]][1]}) %>% unique()
# grep("_NPC_", colnames(GSE70138@mat), value=T) %>% sapply(., function(i) {stringr::str_split(i, ":")[[1]][1]}) %>% unique()
# grep("_MNEU.E_", colnames(GSE70138@mat), value=T) %>% sapply(., function(i) {stringr::str_split(i, ":")[[1]][1]}) %>% unique()
# grep("_NPC.CAS9_", colnames(GSE70138@mat), value=T) %>% sapply(., function(i) {stringr::str_split(i, ":")[[1]][1]}) %>% unique()
# grep("_NPC.TAK_", colnames(GSE70138@mat), value=T) %>% sapply(., function(i) {stringr::str_split(i, ":")[[1]][1]}) %>% unique()
# 
# 
# # [1] "LJP007_NEU_24H" "LJP008_NEU_24H" "LJP009_NEU_24H"
# # [1] "LJP007_NPC_24H"  "LJP008_NPC_24H"  "LJP009_NPC_24H"  "LPROT003_NPC_6H"
# # [1] "LJP007_MNEU.E_24H"
# # [1] "LJP007_NPC.CAS9_24H" "LJP008_NPC.CAS9_24H" "LJP009_NPC.CAS9_24H"
# # [1] "LJP007_NPC.TAK_24H"  "LJP008_NPC.TAK_24H"  "LJP009_NPC.TAK_24H"  "LPROT001_NPC.TAK_6H" "LPROT002_NPC.TAK_6H"
```


```{python}
# # Check the data
# sig1 = pd.read_table(data_dir+"GSE92742_Broad_LINCS_sig_info.txt", sep="\t", low_memory=False)
# sig2 = pd.read_table(data_dir+"GSE70138_Broad_LINCS_sig_info.txt", sep="\t", low_memory=False)
# sig1_fil = sig1.loc[(sig1["cell_id"]=="NEU") | (sig1["cell_id"]=="NPC") | (sig1["cell_id"]=="MNEU.E") | (sig1["cell_id"]=="NPC.CAS9") | (sig1["cell_id"]=="NPC.TAK")]
# sig2_fil = sig2.loc[(sig2["cell_id"]=="NEU") | (sig2["cell_id"]=="NPC") | (sig2["cell_id"]=="MNEU.E") | (sig2["cell_id"]=="NPC.CAS9") | (sig2["cell_id"]=="NPC.TAK")]
# 
# sorted(set(sig1_fil.iloc[:, 7].to_list()).union(set(sig2_fil.iloc[:, 5].to_list())))
# # ['-666', '0.04 um', '0.05 um', '0.1 %', '0.1 um', '0.12 um', '0.25 um', '0.37 um', '0.4 um', '0.5 um', '0.75 um', '1 µM', '1.0 um', '1.11 um', '1.5 um', '1.5 µL', '1.7 um', '10 µM', '10.0 um', '100 nM', '12.0 um', '2.0 um', '2.5 um', '25.0 um', '3 µM', '3.0 um', '3.33 um', '3.5 um', '4.0 um', '4.5 um', '40.0 um', '5.0 um', '5.5 um', '500 nM', '6.0 um', '8.0 um']
# ['0.04 um' '0.05 um' '0.1 um' '0.12 um' '0.25 um' '0.37 um' '0.4 um' '0.5 um' '0.75 um' '1.0 um' '1.11 um' '1.5 um' '1.7 um' '10.0 um' '100.0 um' '12.0 um' '2.0 um' '25.0 um' '3.0 um' '3.33 um' '3.5 um' '4.0 um' '4.5 um' '40.0 um' '5.0 um' '5.5 um' '500.0 nm' '6.0 um' '8.0 um']
# ['0.04 µM' '0.05 µM' '0.1 µM' '0.12 µM' '0.25 µM' '0.37 µM' '0.4 µM' '0.5 µM' '0.75 µM' '1 µM' '1.11 µM' '1.5 µM' '1.7 µM' '10 µM' '100 µM' '12 µM' '2 µM' '25 µM' '3 µM' '3.33 µM' '3.5 µM' '4 µM' '4.5 µM' '40 µM' '5 µM' '5.5 µM' '500 nM' '6 µM' '8 µM']
# sorted(set(sig1_fil.iloc[:, 10].to_list()).union(set(sig2_fil.iloc[:, 6].to_list())))
# # ['24 h', '6 h', '96 h']
# 
# 
# ctrl1 = sig1_fil.loc[(sig1_fil["pert_type"]=='ctl_vector.cns') | (sig1_fil["pert_type"]=="ctl_vehicle.cns") | (sig1_fil["pert_type"]=="ctl_untrt.cns") | (sig1_fil["pert_type"]=="ctl_vehicle") | (sig1_fil["pert_type"]=="ctl_untrt")  | (sig1_fil["pert_type"]=="ctl_vector")]
# case1 = sig1_fil.loc[(sig1_fil["pert_type"]=='trt_sh.cgs') | (sig1_fil["pert_type"]=='trt_sh.css') | (sig1_fil["pert_type"]=='trt_cp') | (sig1_fil["pert_type"]=='trt_sh')]
# set(ctrl1.iloc[:, 7].to_list())
# # {'-666', '1.5 µL', '0.1 %'}  # It's good that we don't use those doses when filtering the data, only used those for treatment samples.
# set(case1.iloc[:, 7].to_list())
# # {'3 µM', '1 µM', '100 nM', '1.5 µL', '500 nM', '10 µM'}
# set(ctrl1.iloc[:, 10].to_list())
# set(case1.iloc[:, 10].to_list())
# set(case1.iloc[:, 3].to_list())
# # {'trt_sh.cgs', 'trt_sh', 'trt_sh.css', 'trt_cp'}
# 
# ctrl2 = sig2_fil.loc[(sig2_fil["pert_type"]=='ctl_vector.cns') | (sig2_fil["pert_type"]=="ctl_vehicle.cns") | (sig2_fil["pert_type"]=="ctl_untrt.cns") | (sig2_fil["pert_type"]=="ctl_vehicle") | (sig2_fil["pert_type"]=="ctl_untrt")  | (sig2_fil["pert_type"]=="ctl_vector")]
# case2 = sig2_fil.loc[(sig2_fil["pert_type"]=='trt_sh.cgs') | (sig2_fil["pert_type"]=='trt_sh.css') | (sig2_fil["pert_type"]=='trt_cp') | (sig2_fil["pert_type"]=='trt_sh')]
# set(ctrl2.iloc[:, 5].to_list())
# # {'-666'} # It's good that we don't use those doses when filtering the data, only used those for treatment samples.
# set(case2.iloc[:, 5].to_list())
# # {'2.5 um', '5.5 um', '0.37 um', '0.25 um', '0.5 um', '10.0 um', '0.4 um', '1.7 um', '12.0 um', '8.0 um', '1.5 um', '3.5 um', '4.0 um', '6.0 um', '3.33 um', '25.0 um', '0.12 um', '0.04 um', '0.75 um', '1.11 um', '4.5 um', '0.05 um', '1.0 um', '5.0 um', '2.0 um', '3.0 um', '0.1 um', '40.0 um'}
# set(ctrl2.iloc[:, 6].to_list())
# set(case2.iloc[:, 6].to_list())
# set(case2.iloc[:, 3].to_list())
# # {'trt_cp'}
```

```{bash}
# Extract all the perturbation signature data for cell lines NEU, MNEU.E, NPC, NPC.CAS9, NPC.TAK from L1000
# bash ../results/rerun/Additional/DrugTarget_L1000/automation.sh
```


```{r}
# sig1 = read.table("../results/rerun/Additional/DrugTarget_L1000/data/GSE92742_Broad_LINCS_sig_info.txt", 
#                   header = T, quote = "", sep="\t") %>% as.data.frame()
# sig2 = read.table("../results/rerun/Additional/DrugTarget_L1000/data/GSE70138_Broad_LINCS_sig_info.txt", 
#                   header = T, quote = "", sep="\t") %>% as.data.frame()
# filter(sig1, pert_type=="trt_sh")

# Prepare for gene name conversion
genes = read.table("../results/rerun/Additional/DrugTarget_L1000/data/GSE92742_Broad_LINCS_gene_info.txt", 
                  header = T, quote = "", sep="\t") %>% as.data.frame()
mappings = setNames(genes$pr_gene_symbol, genes$pr_gene_id)

# The perturbation signatures: when there are replicates for one combination (e.g. MNEU.E_24 h_0.05 µM, cell line: MNEU.E, pert_time: 24 h, pert_dose: 0.05 uM), the expression signature from all replicates are combined (median)
Names_C = list.files("../results/rerun/Additional/DrugTarget_L1000/Consensus_Signatures/", pattern=".txt") %>% 
  gsub("_Consensus.txt", "", .)
Consensus = list.files("../results/rerun/Additional/DrugTarget_L1000/Consensus_Signatures/", pattern=".txt", 
                       full.names=T) %>% setNames(Names_C)
length(Consensus)
length(Names_C)

# The perturbation signatures: when there are no replicates
Names_N = list.files("../results/rerun/Additional/DrugTarget_L1000/Compounds_With_No_Replicates/", pattern=".txt") %>% 
  gsub("_No_Replicates.txt", "", .)
NoReplicates = list.files("../results/rerun/Additional/DrugTarget_L1000/Compounds_With_No_Replicates/", pattern=".txt", 
                          full.names=T) %>% setNames(Names_N)

length(NoReplicates)
length(Names_N)
setdiff(Names_N, Names_C)
Names_all = c(Names_C, Names_N) %>% unique()
length(Names_all)

# Combine all the consensus signatures and non-replicated signatures together into one single table for each dose+time+cellline combination:
mat_all = lapply(Names_all, function(Name) {
  if (Name %in% Names_C) {
    fileC = Consensus[Name]
    matC = read.table(fileC, header = T, quote = "", sep="\t") %>%
      as.data.frame()
  } else matC=NA
  if (Name %in% Names_N) {
    fileN = NoReplicates[Name]
    matN = read.table(fileN, header = T, quote = "", sep="\t") %>%
      as.data.frame()
  } else matN=NA
  if (!is.data.frame(matC)&&is.data.frame(matN)) mat=matN
  if (!is.data.frame(matN)&&is.data.frame(matC)) mat=matC
  if (is.data.frame(matN)&&is.data.frame(matC)) {
    mat = rbind(matC, matN)
    # print(setdiff(colnames(matC),colnames(matN)))
    # print(colnames(matC))
    # print(colnames(matN))
    }
  if (nrow(mat)!=0) {
    colnames(mat) = gsub("X", "", colnames(mat))
    mat = mat %>% tibble::column_to_rownames(var="Compound_id")
    colnames(mat) = mappings[colnames(mat)] %>% as.character()
    mat
  }
}) %>% setNames(Names_all)

# Remove empty items
mat_all = mat_all %>% Filter(Negate(is.null), .)
# Check the number of pertubagens
lapply(mat_all, function(mat) dim(mat))

# How many distinct chemical compounds do we have:
lapply(mat_all, function(m) {
  rownames(m)
}) %>% unlist() %>% as.character() %>% unique() %>% length()
# [1] 3497

# Import logFC information from our Limma differential analysis
outDir = "../results/rerun/"
OUTDIR_QC = paste0(outDir, "/QC/")
dataTable.list_SL_IRS_TMM =
  readRDS(paste0(OUTDIR_QC, "/NormalizedIntensity_dataTable.list_SL_IRS_TMM.RDS"))
NAMES = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
OUTDIR_DA = paste0(outDir, "/DifferentialAnalysis/")
rslt.list = lapply(names(dataTable.list_SL_IRS_TMM), function(i) {
  read_excel(paste0(OUTDIR_DA, "/Differential_Table_", i, ".xlsx"),
                 sheet=1, col_names=TRUE, na="", skip=0)
}) %>% setNames(NAMES)
rslt.list
length(mat_all)

saveRDS(mat_all, "../results/rerun/Additional/DrugTarget_L1000/mat_all.RDS")
# mat_all = readRDS("../results/rerun/Additional/DrugTarget_L1000/mat_all.RDS")
```

## Use all drug signaturess

```{r}
# Calculate Pearson correlation between limma logFC (3 fractions separately) and the in vitro L1000 compound signatures
drugRepurposing = function(dat.list=rslt.list, mat_all) {
  NAMES = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
  # The list below has three layers: cellline+dose+time combination—>cellular fractions—>perturbagens
  cor_ests = lapply(mat_all, function(l) {
    lapply(dat.list, function(d) {
      shared_genes = intersect(colnames(l), d$Genes)
      l = l[, shared_genes]
      d_mappings = setNames(d$logFC, d$Genes)
      logFC = d_mappings[shared_genes]
      cors = sapply(rownames(l), function(i) {
        x = l[i, shared_genes] %>% as.numeric()
        y = logFC[shared_genes] %>% as.numeric()
        cor(x, y, method = "pearson")
      }) %>% setNames(rownames(l))
    }) %>% setNames(NAMES)
  }) %>% setNames(names(mat_all))
  
  # Combine the perturbagens for each cellular fraction into one single table:
  cor_ests2 = lapply(names(cor_ests), function(i) {
    c = cor_ests[[i]]
    lapply(names(c), function(j) {
      c[[j]]
    }) %>% as.data.frame() %>% 
      setNames(NAMES)
  }) %>% setNames(names(mat_all))
  
  shared_genes = intersect(colnames(mat_all[[1]]), dat.list[[1]]$Genes)
  
  # cellines = c("NEU", "NPC")
  # Bring cellular fraction to the out layer of the list:
  cor_ests3 = lapply(NAMES, function(f) {
    #  lapply(cellines, function(c) {
    #combi = grep(c, names(cor_ests2), value=T)
    combi = names(cor_ests2)
    lapply(seq_along(cor_ests2[combi]), function(i) {
      c=cor_ests2[combi][[i]]
      N=names(cor_ests2[combi])[i]
      mat = c[, f, drop=F] %>% setNames(N)
      mat = mat %>% tibble::rownames_to_column(var="perturbants")
      mat
    }) %>% Reduce(function(x, y) merge(x, y, by="perturbants", all=T), .)    
    # }) %>% setNames(cellines)
  }) %>% setNames(NAMES)
  
  # show the sample size for each correlation calculation
  sample_size = lapply(mat_all, function(l) {
    lapply(dat.list, function(d) {
      shared_genes = intersect(colnames(l), d$Genes)
      length(shared_genes)
    }) %>% setNames(NAMES) %>% as.data.frame() %>% t()
  }) %>% setNames(names(mat_all)) %>% as.data.frame()
  
  Non_zeros_index = which((sample_size[,1, drop=F] %>% t()) != 0)
  
  # In the list below: the out layer are cellular fractions
  metas = lapply(Non_zeros_index, function(j) {
    c = cor_ests3[[j]]
    #lapply(f, function(c) {
    c = c %>% tibble::column_to_rownames(var="perturbants")
    meta = lapply(1:nrow(c), function(i) {
      r = c[i, ,drop=F] %>% t() %>% na.omit() %>% t() %>% as.numeric()
      n = rep(sample_size[,1][j])
      # Calculate Fisher's Z transformation of correlations
      #if (length(r)>1) {
      z <- atanh(r) # Fisher's Z transformation
      v <- 1 / (n - 3) # variances of Fisher's Z
      # Perform the meta-analysis with weights as the inverse of the variances
      res <- rma(yi=z, vi=v, method="FE")
      # Back-transform the pooled estimate to the correlation scale
      pooled_r <- tanh(res$b)
      # Include 95% CI for the pooled correlation coefficient
      ci.lb <- tanh(res$ci.lb)
      ci.ub <- tanh(res$ci.ub)
      col = vector("numeric", length = 0)
      dat = data.frame(cbind(col, col, col, col)) %>% 
        setNames(c("Pooled.cor", "ci.lb", "ci.ub", "p.value"))
      dat[1, ] = c(pooled_r, ci.lb, ci.ub, res$pval)
      rownames(dat) = rownames(c)[i]
      dat
      #}
      #}) %>% setNames(rownames(c)) %>% Filter(Negate(is.null), .) %>% Reduce(rbind, .)
      #meta$adj.p.value = p.adjust(meta$p.value, method = "BH")
      #meta %>% filter(adj.p.value<0.2)
    }) %>% setNames(rownames(c)) %>% Filter(Negate(is.null), .) %>% Reduce(rbind, .)
    meta$adj.p.value = p.adjust(meta$p.value, method = "BH") 
    meta %>% filter(adj.p.value<0.2) # Only keep those with adj.p.value<0.2
  }) %>% setNames(names(cor_ests3)[Non_zeros_index])
  
  # Only keep those with adj.p.value<0.05
  metas = lapply(metas, function(m) {
    filter(m, adj.p.value<0.05)
  })
  return(list(results=metas, genes_used=shared_genes))
}
```


```{r}
# Using all genes
metas = drugRepurposing(dat.list=rslt.list, mat_all)
saveRDS(metas$results, "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates.RDS")
# the p-value for the test of the overall mean effect, which in a two-sided framework, tests null hypothesis that the mean effect is zero against the alternative hypothesis that it is not zer0
# For the interpretation of the result, we need to check the *_Broad_LINCS_sig_info.txt files, to check what pert_type it is 
  ## for trt_cp - Compound (Chemical perturbations), one may be interested to look for perturbagens with Pooled.cor<0

# Save the result to a xlsx file:
tosave = lapply(seq_along(metas$results), function(i){
  m = metas$results[[i]]
  m$Fraction = names(metas$results)[i]
  m = tibble::rownames_to_column(m, var="Perturbagen")
  m$Genes_used = paste(metas$genes_used, collapse = "/")
  m
}) %>% Reduce(rbind, .)
xlsx::write.xlsx(as.data.frame(tosave), 
             "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates.xlsx",
             col.names=TRUE, row.names=FALSE,
             append=F)

# How many show negative correlation:
filter(tosave, Pooled.cor<0) %>% nrow()
# [1] 40

# How many per fraction:
table(tosave$Fraction)
   # CYTOSOLIC    INSOLUBLE SYNAPTOSOMAL 
   #         2           27           15 

filter(tosave, p.value<0.0000001)
```


```{r}
# Using only those significant differential genes from Limma test
# Only keep the differential genes
rslt.list_sig = lapply(rslt.list, function(r) {
  as.data.frame(r) %>% filter(adj.P.Val<0.2)
})

metas_sig = drugRepurposing(dat.list=rslt.list_sig, mat_all)
saveRDS(metas_sig$results, "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates_onlySignificantDEGs.RDS")
# the p-value for the test of the overall mean effect, which in a two-sided framework, tests null hypothesis that the mean effect is zero against the alternative hypothesis that it is not zer0
# For the interpretation of the result, we need to check the *_Broad_LINCS_sig_info.txt files, to check what pert_type it is 
  ## for trt_cp - Compound (Chemical perturbations), one may be interested to look for perturbagens with Pooled.cor<0

# Save the result to a xlsx file:
metas_sig$results = lapply(metas_sig$results, function(meta) {
  if (nrow(meta)==0) {
    NULL
  } else {
    meta
  }}) %>% Filter(Negate(is.null), .)

tosave = lapply(seq_along(metas_sig$results), function(i){
  m = metas_sig$results[[i]]
  m$Fraction = names(metas_sig$results)[i]
  m = tibble::rownames_to_column(m, var="Perturbagen")
  m$Genes_used = paste(metas_sig$genes_used, collapse = "/")
  m
}) %>% Reduce(rbind, .)
xlsx::write.xlsx(as.data.frame(tosave), 
             "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates_onlySignificantDEGs.xlsx",
             col.names=TRUE, row.names=FALSE,
             append=F)

# How many show negative correlation:
filter(tosave, Pooled.cor<0) %>% nrow()
# [1] 0

# How many per fraction:
table(tosave$Fraction)
   # CYTOSOLIC    
   #         2          

filter(tosave, p.value<0.0000001)
```


## Use only Strong drug signatures (>2|<-2)

```{r}
# Calculate Pearson correlation between limma logFC (3 fractions separately) and the in vitro L1000 compound signatures
drugRepurposing_strong = function(dat.list=rslt.list, mat_all) {
  NAMES = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
  # The list below has three layers: cellline+dose+time combination—>cellular fractions—>perturbagens
  cor_ests = lapply(mat_all, function(l) {
    lapply(dat.list, function(d) {
      shared_genes = intersect(colnames(l), d$Genes)
      l = l[, shared_genes]
      d_mappings = setNames(d$logFC, d$Genes)
      logFC = d_mappings[shared_genes]
      cors = sapply(rownames(l), function(i) {
        x = l[i, shared_genes] %>% as.numeric()
        index_strong = (x>2)|(x< -2)
        if (sum(index_strong)>3) {
          x = x[index_strong]
          y = logFC[shared_genes] %>% as.numeric()
          y = y[index_strong]
          cor(x, y, method = "pearson")
        }}) %>% setNames(rownames(l))
      used_genes = sapply(rownames(l), function(i) {
        x = l[i, shared_genes] %>% as.numeric()
        index_strong = (x>2)|(x< -2)
        if (sum(index_strong)>3) {
          shared_genes[index_strong]
        }}) %>% setNames(rownames(l))
      list(cors = cors, used_genes = used_genes)
    }) %>% setNames(NAMES)
  }) %>% setNames(names(mat_all))
  
  # Combine the perturbagens for each cellular fraction into one single table:
  cor_ests2 = lapply(names(cor_ests), function(i) {
    c = cor_ests[[i]]
    lapply(names(c), function(j) {
      c[[j]]$cors %>% unlist() %>% data.frame() %>% setNames(j) %>%
        tibble::rownames_to_column(var="perturbants")
    }) %>% Reduce(full_join, .) %>% 
      tibble::column_to_rownames(var="perturbants")
  }) %>% setNames(names(mat_all))
  
  # cellines = c("NEU", "NPC")
  # Bring cellular fraction to the out layer of the list:
  cor_ests3 = lapply(NAMES, function(f) {
    #  lapply(cellines, function(c) {
    #combi = grep(c, names(cor_ests2), value=T)
    combi = names(cor_ests2)
    lapply(seq_along(cor_ests2[combi]), function(i) {
      c=cor_ests2[combi][[i]]
      N=names(cor_ests2[combi])[i]
      mat = c[, f, drop=F] %>% setNames(N)
      mat = mat %>% tibble::rownames_to_column(var="perturbants")
      mat
    }) %>% Reduce(function(x, y) merge(x, y, by="perturbants", all=T), .)    
    # }) %>% setNames(cellines)
  }) %>% setNames(NAMES)
  
  # Check how many drug disturbance studies/combinations involved for each durg:
  lapply(cor_ests3, function(i) {apply(i, 1, function(j) {sum(!is.na(j))})})
  # Mostly 2-5
  saveRDS(cor_ests3, "../results/rerun/Additional/DrugTarget_L1000/Correlations_single_studies.RDS")
  
  # show the sample size for each correlation calculation
  sample_size = lapply(dat.list, function(d) {
    lapply(mat_all, function(l) {
      shared_genes = intersect(colnames(l), d$Genes)
      sapply(rownames(l), function(i) {
        x = l[i, shared_genes] %>% as.numeric()
        index_strong = (x>2)|(x< -2)
        if (sum(index_strong)>3) {
          sum(index_strong)
        }
      }) %>% setNames(rownames(l))
    }) %>% setNames(names(mat_all)) #%>% as.data.frame() %>% t()
  }) %>% setNames(NAMES) #%>% as.data.frame()
  
  sample_size2 = lapply(sample_size, function(fraction) {
      Ns = names(fraction)
      lapply(seq_along(fraction), function(k) {
        rep = fraction[[k]] %>% unlist() %>% 
          data.frame() %>% setNames(Ns[k]) %>%
          tibble::rownames_to_column(var="perturbants")
      }) %>% Reduce(full_join, .) %>%
        tibble::column_to_rownames(var="perturbants")
    }) %>% setNames(NAMES) 
  
  # Check the average number of genes that have been used:
  lapply(sample_size2, function(s) {apply(s, 2, function(j) mean(j[!is.na(j)])) %>% mean()} )
  # 31-36
  saveRDS(sample_size2, "../results/rerun/Additional/DrugTarget_L1000/Sample_sizes_single_studies.RDS")
  
  used_genes = lapply(NAMES, function(j) {
    lapply(names(cor_ests), function(i) {
      print(i)
      c = cor_ests[[i]]
      if (sum(!is.na(names(c[[j]]$used_genes)))>1) {
        c[[j]]$used_genes %>% sapply(., function(gl) paste0(gl, collapse = "/")) %>% 
          unlist() %>% data.frame() %>% setNames(i) %>%
          tibble::rownames_to_column(var="perturbants")
      } else if (sum(!is.na(names(c[[j]]$used_genes)))==1) {
        temp = c[[j]]$used_genes[,1] %>% 
          paste0(., collapse = "/") %>%
          data.frame() %>% setNames(i)
        rownames(temp) = names(c[[j]]$used_genes)[!is.na(names(c[[j]]$used_genes))]
        temp %>% tibble::rownames_to_column(var="perturbants")
      }
    }) %>% Reduce(full_join, .) %>% 
      tibble::column_to_rownames(var="perturbants")
  }) %>% setNames(NAMES)
  
  # In the list below: the out layer are cellular fractions
  metas = lapply(seq_along(sample_size2), function(j) {
    c = cor_ests3[[j]] %>% tibble::column_to_rownames(var="perturbants")
    c = c[rownames(sample_size2[[j]]), colnames(sample_size2[[j]])]
    meta = lapply(1:nrow(c), function(i) {
      Index_1sample = !is.na(c[i,])
      r = c[i,][Index_1sample] 
      n = sample_size2[[j]][i,][Index_1sample] 
      # We need at least three data points to calculate a correlation coefficient (since with two points, the correlation will always be exactly 1 or -1, and thus not informative)
      keep = n>3
      if (sum(keep) > 0) {
        n = n[keep]
        r = r[keep]
        # Calculate Fisher's Z transformation of correlations
        #if (length(r)>1) {
        z <- atanh(r) # Fisher's Z transformation
        v <- 1 / (n - 3) # variances of Fisher's Z
        # Perform the meta-analysis with weights as the inverse of the variances
        res <- rma(yi=z, vi=v, method="REML")
        # Back-transform the pooled estimate to the correlation scale
        pooled_r <- tanh(res$b)
        # Include 95% CI for the pooled correlation coefficient
        ci.lb <- tanh(res$ci.lb)
        ci.ub <- tanh(res$ci.ub)
        col = vector("numeric", length = 0)
        dat = data.frame(cbind(col, col, col, col)) %>% 
          setNames(c("Pooled.cor", "ci.lb", "ci.ub", "p.value"))
        dat[1, ] = c(pooled_r, ci.lb, ci.ub, res$pval)
        rownames(dat) = rownames(c)[i]
        dat
      }
      #}
      #}) %>% setNames(rownames(c)) %>% Filter(Negate(is.null), .) %>% Reduce(rbind, .)
      #meta$adj.p.value = p.adjust(meta$p.value, method = "BH")
      #meta %>% filter(adj.p.value<0.2)
    }) %>% setNames(rownames(c)) %>% Filter(Negate(is.null), .) %>% Reduce(rbind, .)
    meta$adj.p.value = p.adjust(meta$p.value, method = "BH") 
    meta %>% filter(adj.p.value<0.2) # Only keep those with adj.p.value<0.2
  }) %>% setNames(NAMES)
  
  # Only keep those with adj.p.value<0.05
  metas = lapply(metas, function(m) {
    filter(m, adj.p.value<0.05)
  })
  return(list(results=metas, genes_used=used_genes))
}
```
### Use all proteins (not only differential proteins)

```{r}
# Using all genes
metas_strong = drugRepurposing_strong(dat.list=rslt.list, mat_all)
saveRDS(metas_strong, "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates_OnlyStrongDrugSignature.RDS")
# the p-value for the test of the overall mean effect, which in a two-sided framework, tests null hypothesis that the mean effect is zero against the alternative hypothesis that it is not zer0
# For the interpretation of the result, we need to check the *_Broad_LINCS_sig_info.txt files, to check what pert_type it is 
  ## for trt_cp - Compound (Chemical perturbations), one may be interested to look for perturbagens with Pooled.cor<0

# Save the result to a xlsx file:
tosave = lapply(seq_along(metas_strong$results), function(i){
  m = metas_strong$results[[i]]
  m$Fraction = names(metas_strong$results)[i]
  m = tibble::rownames_to_column(m, var="Perturbagen")
  #m$Genes_used = paste(metas$genes_used, collapse = "/")
  m
}) %>% Reduce(rbind, .)
xlsx::write.xlsx(as.data.frame(tosave), 
             "../results/rerun/Additional/DrugTarget_L1000/perturbagen_candidates_OnlyStrongDrugSignatures.xlsx",
             col.names=TRUE, row.names=FALSE,
             append=F)

tosave_usedGenes = lapply(seq_along(metas_strong$genes_used), function(i){
  m = metas_strong$genes_used[[i]]
  m$Fraction = names(metas_strong$genes_used)[i]
  m = tibble::rownames_to_column(m, var="Perturbagen")
  #m$Genes_used = paste(metas$genes_used, collapse = "/")
  m
}) %>% Reduce(rbind, .)
xlsx::write.xlsx(as.data.frame(tosave_usedGenes), 
             "../results/rerun/Additional/DrugTarget_L1000/UsedGenes4Identifying_perturbagen_candidates_OnlyStrongDrugSignatures.xlsx",
             col.names=TRUE, row.names=FALSE,
             append=F)

# # How many show negative correlation:
# filter(tosave, Pooled.cor<0) %>% nrow()
# # [1] 40
# 
# # How many per fraction:
# table(tosave$Fraction)
#    # CYTOSOLIC    INSOLUBLE SYNAPTOSOMAL 
#    #         2           27           15 
# 
# filter(tosave, p.value<0.0000001)
```


# Using Phosphoproteomics drug signatures from LINCS p100 data

```{r}
library(dplyr)
RT = data.table::fread("../results/rerun/Additional/DrugTarget_L1000/P100/P100-aggregated-data-Chemical-perturbations-Level4.2020_feb.processed.gct", sep="\t", header=T)

NPC = as.data.frame(RT)[, c(1:20,which(RT[1,]=="NPC"))]

# There are 4 plates: [1] "P-0027" "P-0016" "P-0020" "P-0063"
plates = NPC[4,] %>% as.character() %>% unique()
# Within each plate, substract the average log2 DMSO ratio (level 4 data), and then take z score per sample
all = lapply(plates, function(p) {
  proteins = NPC$pr_uniprot_id[25:nrow(NPC)]
  duplicated_proteins = proteins[duplicated(proteins)]
  for (protein in duplicated_proteins) {
    combined = NPC[NPC$pr_uniprot_id==protein, 21:ncol(NPC)] %>%
      apply(., 2, function(c) as.numeric(c)) %>% 
      2^. %>% colSums(., na.rm=T) %>% log2(.) 
    NPC[NPC$pr_uniprot_id==protein, 21:ncol(NPC)][1,][combined!=0] =
    combined[combined!=0]
  }
  NPC = rbind(NPC[1:24, ], NPC[25:nrow(NPC),][!duplicated(NPC$pr_uniprot_id[25:nrow(NPC)]),])
  proteins_unique = NPC$pr_uniprot_id[25:nrow(NPC)]
  # Calcualte the average log2 DMSO ratio
  Ctrl = NPC[25:nrow(NPC), NPC[14,]=="DMSO"&NPC[4,]==p] %>%   apply(., 2, function(c) as.numeric(c)) %>% rowMeans(., na.rm=T)
  Others = NPC[25:nrow(NPC), NPC[14,]!="DMSO"&NPC[4,]==p] %>% apply(., 2, function(c) as.numeric(c))
  # substract the average log2 DMSO ratio (level 4 data), and then take z score per sample
  trans.diff = (Others-Ctrl) %>% apply(., 2, function(o) {
  (o-mean(o, , na.rm=T))/sd(o, na.rm=T)})
  pert_iname = NPC[14,NPC[14,]!="DMSO"&NPC[4,]==p] %>% as.character() %>% unique()
  iname = lapply(pert_iname, function(i) {
    i_index = (NPC[14,NPC[14,]!="DMSO"&NPC[4,]==p] %>% as.character())==i
    temp = trans.diff[, i_index]  %>% rowMeans(., na.rm=T) %>% as.data.frame(list(.)) %>% setNames(i) %>% as.data.frame() %>% t() 
    colnames(temp) = proteins_unique
    temp
  }) %>% Reduce(rbind, .)
  temp = iname[, colSums(iname, na.rm=T)!=0] %>% as.data.frame() %>% tibble::rownames_to_column(var="Compound") %>%
    dplyr::select(., Compound, everything())
  temp
}) %>% setNames(plates)

# Import logFC information from our Limma differential analysis
outDir = "../results/rerun/"
OUTDIR_QC = paste0(outDir, "/QC/")
dataTable.list_SL_IRS_TMM =
  readRDS(paste0(OUTDIR_QC, "/NormalizedIntensity_dataTable.list_SL_IRS_TMM.RDS"))
NAMES = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
OUTDIR_DA = paste0(outDir, "/DifferentialAnalysis/")
rslt.list = lapply(names(dataTable.list_SL_IRS_TMM), function(i) {
  read_excel(paste0(OUTDIR_DA, "/Differential_Table_", i, ".xlsx"),
                 sheet=1, col_names=TRUE, na="", skip=0)
}) %>% setNames(NAMES)


# Calculate correlation
cor_ests = lapply(rslt.list, function(d) {
    lapply(all, function(l) {
      l = as.data.frame(l) %>%
    tibble::column_to_rownames(var="Compound")
      shared_features = intersect(colnames(l), d$Protein_IDs)
      l = l[, shared_features]
      d_mappings = setNames(d$logFC, d$Protein_IDs)
      logFC = d_mappings[shared_features]
      cors = sapply(rownames(l), function(i) {
        x = l[i, shared_features] %>% as.numeric()
        index_nonNaN = !is.na(x)
        if (sum(index_nonNaN)>3) {
          x = x[index_nonNaN]
          y = logFC[shared_features] %>% as.numeric()
          y = y[index_nonNaN]
          rslt = cor.test(x, y, method = "pearson")
          list(cors=rslt$estimate, p.value= rslt$p.value)
        }}) %>% setNames(rownames(l))
      used_features = sapply(rownames(l), function(i) {
        x = l[i, shared_features] %>% as.numeric()
        index_nonNaN = !is.na(x)
        if (sum(index_nonNaN)>3) {
          shared_features[index_nonNaN]
        }}) %>% setNames(rownames(l))
      list(cors = cors, used_genes = used_features)
    }) %>% setNames(names(all))
  }) %>% setNames(NAMES)

# Combine the perturbagens for each cellular fraction into one single table:
cor_ests2 = lapply(names(cor_ests), function(i) {
    c = cor_ests[[i]]
    cor.table = lapply(names(c), function(j) {
      c[[j]]$cors %>% as.data.frame() %>% t()
    }) %>% Reduce(rbind, .) %>% as.data.frame() # Combine different plates
    cor.table$adj.p = p.adjust(cor.table$p.value, method="fdr")
    cor.table$Fraction = i
    cor.table %>% filter(adj.p<0.05)
  }) %>% setNames(NAMES) %>% Reduce(rbind, .)

dir.create("../results/rerun/Additional/DrugTarget_L1000/P100",
  recursive=T, showWarnings=F)
xlsx::write.xlsx(as.data.frame(cor_ests2), 
             "../results/rerun/Additional/DrugTarget_L1000/P100/sig.correlated.drugs.xlsx",
             col.names=TRUE, row.names=FALSE,
             append=F)

```




# Session Info

```{r}
sink(paste0("../results/rerun/Additional/DrugTarget_L1000/sessionInfo.txt"))
print(sessionInfo())
sink()
```

# Deliverables

# Directory

## Practical info

## Software

# Data responsibility

The responsibility for data archiving lies wscale = TRUEith the PI of the project. We do not offer long-term storage or retrieval of data.

-   **NBIS & UPPNEX:** We kindly ask that you remove the files from UPPMAX/UPPNEX. The main storage at UPPNEX is optimized for high-speed and parallel access, which makes it expensive and not the right place for longer time archiving. Please consider others by not taking up the expensive space. Please note that UPPMAX is a resource separate from the Bioinformatics Platform, administered by the Swedish National Infrastructure for Computing (SNIC) and SNIC-specifc project rules apply to all projects hosted at UPPMAX.
-   **Sensitive data:** Please note that special considerations may apply to the human-derived legally considered sensitive personal data. These should be handled according to specific laws and regulations as outlined e.g. [here](http://nbis.se/support/human-data.html).
-   **Long-term backup:** We recommend asking your local IT for support with long-term data archiving. The [Data Office](https://www.scilifelab.se/data/) at SciLifeLab may be of help to discuss other options.

# Acknowledgments

If you are presenting the results in a paper, at a workshop or conference, we kindly ask you to acknowledge us.

-   NBIS staff are encouraged to be co-authors when this is merited in accordance to the ethical recommendations for authorship, e.g. [ICMJE recommendations](http://www.icmje.org/recommendations/browse/roles-and-responsibilities/defining-the-role-of-authors-and-contributors.html). If applicable, please include **Name, Surname, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Further Affliations**, as co-author. In other cases, NBIS would be grateful if support by us is acknowledged in publications according to this example:

> "Support by NBIS (National Bioinformatics Infrastructure Sweden) is gratefully acknowledged."

-   **UPPMAX** kindly asks you to [acknowledge UPPMAX and NAISS](https://www.uppmax.uu.se/support/faq/general-miscellaneous-faq/acknowledging-uppmax-and-naiss).

-   **NGI:** For publications based on data from NGI Sweden, NGI and SciLifeLab should be acknowledged. Instructions are included in the NGI reports as described [here](https://ngisweden.scilifelab.se/resources/getting-started-at-ngi/).
