---
title: "Subcellular proteomic analysis of the insula in alcohol dependence"
subtitle: "NBIS support {{< 7314 >}}"
author: Yuan Li
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
  pdf_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
editor_options:
  chunk_output_type: inline
params:
  dataset: 1
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)


```

```{r,echo=FALSE,include=FALSE}
# project variables
rep_nbis_id <- "7314"
rep_report_version <- "1.0"
rep_request <- "Philippe Melas"
rep_request_email <- "philippe.melas@ki.se"
rep_pi <- "Philippe Melas"
rep_pi_email <- "philippe.melas@ki.se"
rep_org <- "KI"
rep_nbis <- "Yuan Li"
rep_nbis_email <- "yuan.li@nbis.se"
```

::: boxy
**NBIS ID:** `r rep_nbis_id`\
**Report Version:** `r rep_report_version`\
**Request by:** `r paste0(rep_request," (",rep_request_email,")")`\
**Principal Investigator:** `r paste0(rep_pi," (",rep_pi_email,")")`\
**Organisation:** `r rep_org`\
**NBIS Staff:** `r paste0(rep_nbis," (",rep_nbis_email,")")` **conda env:** TTN
:::

<br>

# Support Request

[7314](https://projects.nbis.se/issues/7314#change-40330) In brief, the data that we have, derive from the following experimental design: Brain samples from alcohol dependent patients (N=12 cases) and matched controls (N=12 controls) were first submitted to subcellular fractionations, which yielded three fractions (i.e., synaptosomes, cytosol, insoluble fraction). All three fractions (total N=72 samples) were subjected to TMT-based quantitative proteomics. Data were provided by a commercial company (Creative Proteomics) in .xlsx format. The company also performed and provided results from basic bioinformatic comparisons (inc. differential expression, pathway analysis, etc.), which revealed some statistical significances between cases and controls on the nominal p-level. We would now like to re-analyze the data and prepare material, e.g., Figures/Tables and Methods, for a publication.

# Data

The data has already been filtered by the company (e.g. removed proteins with CONTA, REV... as prefix of their names, some **potential** contaminating proteins etc. were kept for they may also represent real proteins)

Because there is a big batch effect (between G1 and G2, see the PCA plot in ../results/old/totalAbundanceNormalized/PELLET/differentialAnalysis/Comparisons_loess0.1/Norm-stats-report-Comparisons_loess0.1.pdf), and there are almost no significant protein. So I decided to renormalize the raw intensity, and see how it will affect the results.

**Aim for this particular notebook**: Protein-Protein Interaction (PPI) network analysis using SynaptomeDB: I also think that a PPI could provide complementary information to the GO analysis since synaptic function is largely regulated by dynamic protein complexes, e.g., receptor clustering and scaffolding proteins (so would be interesting to see if any key hubs and/or clusters emerge). When looking online, I found that SynaptomeDB focuses on synaptic PPI networks, which could be particularly useful for identifying altered interaction hubs in AUD:
https://bioconductor.org/packages/release/data/annotation/vignettes/synaptome.db/inst/doc/synaptome_db_query.html

Do you agree that these would be better analytic approaches for the synaptosomal fraction?
p.s. I know that limma didn't identify any significances in the synaptosomal fraction, so for these analyses we would need to input only the PLS-DA differential proteins (and of course the results would then need to be treated as exploratory).

```{r}
#devtools::install_github('lptolik/synaptome.db')
invisible(capture.output(library(synaptome.db)))
invisible(capture.output(library(dplyr)))
invisible(capture.output(library(igraph)))
```

```{r}
OUTPUT = "../results/rerun/plsDA/synaptome.db/"
```


# Import plsDA selected proteins for synaptosomal

```{r}
plsDA_out_syn = 
  readxl::read_excel("../results/rerun/plsDA/pathwayAnalysis/top_loading_proteins_used4ORA.xlsx") %>% 
  as.data.frame() %>% filter(Fraction=="SYNAPTOSOMAL")
up = filter(plsDA_out_syn, Reg=="up") %>% pull(Genes)
down = filter(plsDA_out_syn, Reg=="down") %>% pull(Genes)
all = c(up, down)
up
down
length(all)
# 107
```

# Check pls-DA selected proteins against synaptic proteom db

```{r}
t = All_HumanEntrez = findGenesByName(all) %>% 
  pull(HumanEntrez) %>% 
  getGeneInfoByEntrez(.)
head(t)

# See how many genes are mapped to different brain regions
# We can see most genes are mapped to Cerebral cortex, where insula is located.
t %>% group_by(BrainRegion, HumanName) %>% 
  summarise(n = n())  %>% pull(BrainRegion) %>%
  table()
```


```{r, fig.height=15, fig.width=15, error=TRUE}
# type of the PPI network should be either induced (for all the PPIs for specific genes, including external genes) or limited (for PPIs between the genes specified in the query). Type could be shortened to recognizable minimum like 'ind' or 'lim'.
g = getIGraphFromPPI(
    getPPIbyName(all, type='induced'))
tbl = getTableFromPPI(getPPIbyName(all, type='induced'))
TableNames = list("Synaptosomal_up"=tbl)
openxlsx::write.xlsx(TableNames, 
                     paste0(OUTPUT,"/PPInetwork.xlsx"))

lapply(seq(10,50,10), function(minEdges) {
  L = (70-minEdges)/25
  label.hubs = c(tbl$HumanName.A, tbl$HumanName.B) %>% table() %>%
    .[.>minEdges] %>% names() %>% intersect(all)
  # Map each vertex to its group's color
  vertex.colors = ifelse(
    igraph::V(g)$name %in% synaptome.db:::getGeneIdByName(up), 
    "red", 
    ifelse(igraph::V(g)$name %in% synaptome.db:::getGeneIdByName(down), "blue", "black"))
  vertex.label = rep("", igraph::V(g)$HumanName%>%length())
  vertex.label[igraph::V(g)$HumanName %in% label.hubs] = igraph::V(g)$HumanName[igraph::V(g)$HumanName %in% label.hubs]
  pdf(paste0(OUTPUT, "/PPInetwork_onlyHubsFromThisStudy_minEdge", minEdges,"_synaptosomal.pdf"), h=10*L, w=10*L)
  plot(g,vertex.label=vertex.label, vertex.frame.color="snow", vertex.shape="none",
       vertex.size=0, #vertex.label.degree = pi/3,
       vertex.color = "snow", edge.color = "snow3",
       vertex.label.color = vertex.colors,
       layout=layout_with_fr(g))
  dev.off() 
  
  pdf(paste0(OUTPUT, "/PPInetwork_synaptosomal.pdf"), h=35, w=30)
  plot(g,vertex.label=igraph::V(g)$HumanName,
     vertex.size=0, vertex.shape="none",
     vertex.label.color = vertex.colors,
     edge.color = "snow2")
  dev.off()
})



```


# Session Info

```{r}
output = capture.output(sessionInfo())
writeLines(output, paste0(OUTPUT, "/sessionInfo.txt"))

```

# Method summary
To investigate the brain cell types that are potentially responsible for AUD, we performed Gene Set Enrichment Analysis (GSEA) using the fgsea R package (v.1.24.0) with default setting. Pre-ranked gene lists were generated based on log2 fold-change from Limma differential expression analysis between AUD and UC.

Enrichment analysis was conducted against brain cell type markers obtained from the DISCO (https://www.immunesinglecell.org/atlas/brain). 

The threshold for significance was set at an FDR-adjusted p-value (padj) of less than 0.05. Enrichment results were summarized with normalized enrichment scores (NES), nominal p-values, and FDR q-values. Leading-edge analysis was conducted to identify the subset of genes that contributed most significantly to the enrichment results for each significantly enriched celltype.

The fgsea analysis revealed significant enrichment of 4 celltypes (Protoplasmic astrocyte, Fibrous astrocyte, Microglia, Endothelial cell) for the cytosolic fraction, 2 celltypes (Endothelial cell, T cell) for the insoluble fraction, but no celltype for the synaptosomal fraction. 

# Deliverables

# Directory

## Practical info

## Software

# Data responsibility

The responsibility for data archiving lies wscale = TRUEith the PI of the project. We do not offer long-term storage or retrieval of data.

-   **NBIS & UPPNEX:** We kindly ask that you remove the files from UPPMAX/UPPNEX. The main storage at UPPNEX is optimized for high-speed and parallel access, which makes it expensive and not the right place for longer time archiving. Please consider others by not taking up the expensive space. Please note that UPPMAX is a resource separate from the Bioinformatics Platform, administered by the Swedish National Infrastructure for Computing (SNIC) and SNIC-specifc project rules apply to all projects hosted at UPPMAX.
-   **Sensitive data:** Please note that special considerations may apply to the human-derived legally considered sensitive personal data. These should be handled according to specific laws and regulations as outlined e.g. [here](http://nbis.se/support/human-data.html).
-   **Long-term backup:** We recommend asking your local IT for support with long-term data archiving. The [Data Office](https://www.scilifelab.se/data/) at SciLifeLab may be of help to discuss other options.

# Acknowledgments

If you are presenting the results in a paper, at a workshop or conference, we kindly ask you to acknowledge us.

-   NBIS staff are encouraged to be co-authors when this is merited in accordance to the ethical recommendations for authorship, e.g. [ICMJE recommendations](http://www.icmje.org/recommendations/browse/roles-and-responsibilities/defining-the-role-of-authors-and-contributors.html). If applicable, please include **Name, Surname, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Further Affliations**, as co-author. In other cases, NBIS would be grateful if support by us is acknowledged in publications according to this example:

> "Support by NBIS (National Bioinformatics Infrastructure Sweden) is gratefully acknowledged."

-   **UPPMAX** kindly asks you to [acknowledge UPPMAX and NAISS](https://www.uppmax.uu.se/support/faq/general-miscellaneous-faq/acknowledging-uppmax-and-naiss).

-   **NGI:** For publications based on data from NGI Sweden, NGI and SciLifeLab should be acknowledged. Instructions are included in the NGI reports as described [here](https://ngisweden.scilifelab.se/resources/getting-started-at-ngi/).
