---
title: "Subcellular proteomic analysis of the insula in alcohol dependence"
subtitle: "NBIS support {{< 7314 >}}"
author: Yuan Li
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
  pdf_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
editor_options:
  chunk_output_type: inline
params:
  dataset: 1
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)


```

```{r,echo=FALSE,include=FALSE}
# project variables
rep_nbis_id <- "7314"
rep_report_version <- "1.0"
rep_request <- "Philippe Melas"
rep_request_email <- "philippe.melas@ki.se"
rep_pi <- "Philippe Melas"
rep_pi_email <- "philippe.melas@ki.se"
rep_org <- "KI"
rep_nbis <- "Yuan Li"
rep_nbis_email <- "yuan.li@nbis.se"
```

::: boxy
**NBIS ID:** `r rep_nbis_id`\
**Report Version:** `r rep_report_version`\
**Request by:** `r paste0(rep_request," (",rep_request_email,")")`\
**Principal Investigator:** `r paste0(rep_pi," (",rep_pi_email,")")`\
**Organisation:** `r rep_org`\
**NBIS Staff:** `r paste0(rep_nbis," (",rep_nbis_email,")")` **conda env:** TTN
:::

<br>

# Support Request

[7314](https://projects.nbis.se/issues/7314#change-40330) In brief, the data that we have, derive from the following experimental design: Brain samples from alcohol dependent patients (N=12 cases) and matched controls (N=12 controls) were first submitted to subcellular fractionations, which yielded three fractions (i.e., synaptosomes, cytosol, insoluble fraction). All three fractions (total N=72 samples) were subjected to TMT-based quantitative proteomics. Data were provided by a commercial company (Creative Proteomics) in .xlsx format. The company also performed and provided results from basic bioinformatic comparisons (inc. differential expression, pathway analysis, etc.), which revealed some statistical significances between cases and controls on the nominal p-level. We would now like to re-analyze the data and prepare material, e.g., Figures/Tables and Methods, for a publication.

# Data

The data has already been filtered by the company (e.g. removed proteins with CONTA, REV... as prefix of their names, some **potential** contaminating proteins etc. were kept for they may also represent real proteins)

Because there is a big batch effect (between G1 and G2, see the PCA plot in ../results/old/totalAbundanceNormalized/PELLET/differentialAnalysis/Comparisons_loess0.1/Norm-stats-report-Comparisons_loess0.1.pdf), and there are almost no significant protein. So I decided to renormalize the raw intensity, and see how it will affect the results.

**Aim**: run plsDA for each cellular compartment, and see if we can separate the case and control samples.

**Pellet = Insoluble = Nuclear**


# Import the data and preparation

```{r}
suppressPackageStartupMessages({
  library(mixOmics) # import the mixOmics library
  library(dplyr)
  library(tidyverse)
  library(ggplot2)
  library(xlsx)
  library(igraph)
  library(cowplot)
  library(purrr)
  library("glmnet")
})
set.seed(123) # for reproducibility, remove for normal use
```

```{r}
OUTDIR = "../results/rerun/plsDA/"
dir.create(OUTDIR, recursive = T, showWarnings = F)
```

```{r}
# Import indepdent variable Y
meta = read.table("../data/design_7314/design_CYTOSOLIC.tsv", header=T) %>%
  filter(., sample!="133C_G2") # "133C_G2" was removed as an outlier in CYTOSOLIC.
  #mutate(sample=substr(sample, 1,4))
#head(meta)
Y = meta$Condition %>% factor(., levels=c("control", "case"))

# Import normalized intensity tables for all cellular compartments together
normData = 
  readRDS("../results/rerun/QC/NormalizedIntensity_dataTable.list_SL_IRS_TMM.RDS")
data.list = lapply(normData, function(n) {
  X = n %>% dplyr::select(-c(Genes, Gene_names)) %>%
    as.data.frame()
  rownames(X) = NULL
  X = X %>% column_to_rownames(var="Protein_IDs") %>% t()
  # rownames(X) = substr(rownames(X), 1,4)
  ordered = setdiff(meta$sample, "133C_G2") 
  # All the compartments should have the same samples
  # remove proteins with NA as well
  X = X[ordered, ] %>% t() %>% na.omit() %>% t() #DIABLO can't handle missing values
})
# In each dataset: samples x proteins:
```

Here, we will skip centering etc because plsda is going to do that for us.


# Modelling

## Build a function for hyperparameter tuning

```{r}
runPlsDA = function(data, Y, folds, nrepeat, OUTDIR, i, sample) {
  ## Construct the original model
  set.seed(123)
  plsda.model = plsda(X=data, Y=Y, ncomp=10)
  
  ## Tuning number of components
  # K fold validation
  set.seed(30) 
  perf.plsda.model <- perf(plsda.model, validation = 'Mfold', folds = folds,
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = nrepeat)         # We suggest nrepeat = 50

  # pdf(paste0(OUTDIR, "/Tune_component_numberpre.pdf"), h=5, w=6)
  par(mfrow=c(1, 1))
  set.seed(123)
  compPlot = plot(perf.plsda.model, sd = TRUE, legend.position = 'horizontal',
                  main = sample)
  # invisible(dev.off())
  ncomp = perf.plsda.model$choice.ncomp["BER", "centroids.dist"]
  ncomp = max(2, ncomp)

  set.seed(123)
  final.plsda.model = plsda(X=data,Y=Y, ncomp = ncomp)
  ## Plotting
  # pdf(paste0(OUTDIR_tune, "/plotIndiv.pdf"), h=7, w=8)
  set.seed(123)
  plsPlot = plotIndiv(final.plsda.model, ind.names = FALSE, legend=TRUE,
          comp=c(1,2), ellipse = FALSE,
          title = paste0(sample, ''),
          X.label = 'PLS-DA comp 1', Y.label = 'PLS-DA comp 2')
  # invisible(dev.off())
  
  ## Performance of the model
  # run repeated CV performance evaluation
  set.seed(123)
  perf.plsda = perf(final.plsda.model, validation = 'Mfold',
                     M = folds, nrepeat = nrepeat,
                     dist = 'centroids.dist')
  set.seed(123)
  background.max <- background.predict(final.plsda.model,
                                     comp.predicted = 2,
                                     dist = 'centroids.dist')
  set.seed(123)
  backgroundPlot = plotIndiv(final.plsda.model, comp = 1:2, group = Y,
          ind.names = FALSE, title = paste0(sample, ''),
          legend = TRUE,  background = background.max)
  
  return(list(final.plsda.model=final.plsda.model, perf.plsda=perf.plsda, 
              compPlot=compPlot, plsPlot=plsPlot, 
              backgroundPlot=backgroundPlot))
}

```

```{r}
folds = 3 # K in K-fold cross validation
nrepeat = 40
compartments = c("CYTOSOLIC", "PELLET", "SYNAPTOSOMAL")
names = c("CYTOSOLIC", "INSOLUBLE", "SYNAPTOSOMAL")
pdf(paste0(OUTDIR, "/plsDA_modeling_plots.pdf"), h=4, w=5)
rslt.list = lapply(seq_along(data.list), function(i) {
  data = data.list[[i]]
  sample = names[i]
  runPlsDA(data, Y, folds, nrepeat, OUTDIR, i, sample) 
}) %>% setNames(compartments)
dev.off()

saveRDS(rslt.list, paste0(OUTDIR, "rslt.list.RDS"))

final.models = lapply(rslt.list, function(r) {
  r$final.plsda.model
})

saveRDS(final.models, paste0(OUTDIR, "final_models.RDS"))
# final.models = readRDS(paste0(OUTDIR, "final_models.RDS"))

sink(paste0(OUTDIR, "final_BER.txt"))
for (r in names(rslt.list)) {
  print(paste0("final plsDA model BER for ", r, ":"))
  print(rslt.list[[r]]$perf.plsda$error.rate$BER)
}
sink()

names = c("CYTOSOLIC", "NUCLEAR", "SYNAPTOSOMAL")
loadings = lapply(seq_along(final.models), function(f) {
  final.models[[f]]$loadings$X %>% as.data.frame() %>% 
    mutate(fraction=names[f]) %>% 
    tibble::rownames_to_column(var="Protein")
}) %>% Reduce(rbind, .)

write.xlsx(loadings, 
             paste0(OUTDIR, "/plsDA_feature_loadings.xlsx"),
             col.names=TRUE, row.names=F,
             append=F)
```


## Top features (|feature loadings|>0.035 with correct logFC direction information) VS differential proteins: upset plot

```{r}
library(readxl)
compartments = c("CYTOSOLIC", "PELLET", "SYNAPTOSOMAL")
rslt.list = lapply(compartments, function(i) {
  read_excel(
    paste0("../results/rerun/DifferentialAnalysis/Differential_Table_",i,".xlsx"), 
                  sheet=1, col_names=TRUE, na="", skip=0)
}) %>% setNames(compartments)
library(ComplexHeatmap)
loading_cutoff = 0.035
makeUpset = function(log2FoldChange_thresh, padj_thresh) {
  protein.list = lapply(names(rslt.list), function(i) {
    up = as.data.frame(rslt.list[[i]]) %>% 
      filter(logFC>log2FoldChange_thresh&adj.P.Val<padj_thresh)
    down = as.data.frame(rslt.list[[i]]) %>% 
      filter(logFC< -log2FoldChange_thresh&adj.P.Val<padj_thresh)
    list(up$Protein_IDs, down$Protein_IDs)
  })  %>% setNames(names(rslt.list))
  
  dir.list = lapply(names(rslt.list), function(i) {
    up = as.data.frame(rslt.list[[i]]) %>% 
      filter(logFC>0)
    down = as.data.frame(rslt.list[[i]]) %>% 
      filter(logFC<0)
    list(up=up$Protein_IDs, down=down$Protein_IDs)
  })  %>% setNames(names(rslt.list))
  
  
  i = 1
  up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.> loading_cutoff] %>% names()
  up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.< -loading_cutoff] %>% names()
  up = c(up1, up2) %>% intersect(dir.list[[i]]$up)
  down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.< -loading_cutoff] %>% names()
  down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.> loading_cutoff] %>% names()
  down = c(down1, down2) %>% intersect(dir.list[[i]]$down)
  C = list(up, down) 
    
  i = 2
  up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.< -loading_cutoff] %>% names()
  up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.> loading_cutoff] %>% names()
  up = c(up1, up2) %>% intersect(dir.list[[i]]$up)
  down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.> loading_cutoff] %>% names()
  down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.< -loading_cutoff] %>% names()
  down = c(down1, down2) %>% intersect(dir.list[[i]]$down)
  I = list(up, down)
  
  i = 3
  up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.< -loading_cutoff] %>% names()
  up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.< -loading_cutoff] %>% names()
  up = c(up1, up2) %>% intersect(dir.list[[i]]$up)
  down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
    .[.> loading_cutoff] %>% names()
  down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
    .[.> loading_cutoff] %>% names()
  down = c(down1, down2) %>% intersect(dir.list[[i]]$down)
  S =list(up, down)
  
  plsDA.features = list(C, I, S) %>% setNames(names(rslt.list))
  compartments_new = c("CYTOSOLIC", "NUCLEAR", "SYNAPTOSOMAL")
  protein.list_UP = c(protein.list[[1]][1], protein.list[[2]][1], 
                      protein.list[[3]][1], plsDA.features[[1]][1],
                      plsDA.features[[2]][1],
                      plsDA.features[[3]][1]) %>% 
    setNames(c(paste0(compartments_new[1], "_UP_Limma"), 
               paste0(compartments_new[2], "_UP_Limma"), 
               paste0(compartments_new[3], "_UP_Limma"),
               paste0(compartments_new[1], "_UP_plsDA"), 
               paste0(compartments_new[2], "_UP_plsDA"), 
               paste0(compartments_new[3], "_UP_plsDA")))
  protein.list_DOWN = c(protein.list[[1]][2], protein.list[[2]][2], 
                        protein.list[[3]][2], plsDA.features[[1]][2],
                      plsDA.features[[2]][2],
                      plsDA.features[[3]][2]) %>% 
    setNames(c(paste0(compartments_new[1], "_DN_Limma"), 
               paste0(compartments_new[2], "_DN_Limma"), 
               paste0(compartments_new[3], "_DN_Limma"),
               paste0(names(protein.list)[1], "_DN_plsDA"), 
               paste0(compartments_new[2], "_DN_plsDA"), 
               paste0(compartments_new[3], "_DN_plsDA")))
  
  # library(ComplexHeatmap)
  pdf(paste0(OUTDIR, "/Sig.Diff._plsDA_top_proteins_upset.pdf"), h=4, w=7)
  m = make_comb_mat(c(protein.list_UP, protein.list_DOWN))
  print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = c(paste0(compartments_new[1], "_UP_Limma"), 
                      paste0(compartments_new[2], "_UP_Limma"), 
                      paste0(compartments_new[3], "_UP_Limma"),
                      paste0(compartments_new[1], "_UP_plsDA"), 
                      paste0(compartments_new[2], "_UP_plsDA"), 
                      paste0(compartments_new[3], "_UP_plsDA"),
                      paste0(compartments_new[1], "_DN_Limma"), 
                      paste0(compartments_new[2], "_DN_Limma"), 
                      paste0(compartments_new[3], "_DN_Limma"),
                      paste0(compartments_new[1], "_DN_plsDA"), 
                      paste0(compartments_new[2], "_DN_plsDA"), 
                      paste0(compartments_new[3], "_DN_plsDA")),
        column_title = paste0("")))
  invisible(dev.off())
  return(list(protein.list_UP=protein.list_UP, protein.list_DOWN=protein.list_DOWN, dir.list=dir.list))
}

log2FoldChange_thresh = 0.3
padj_thresh = 0.2
protein.list = makeUpset(log2FoldChange_thresh, padj_thresh)
saveRDS(protein.list, paste0(OUTDIR, "/Protein.list.RDS"))
```


# Functional pathway analysis

* Here, we combine top (feature loading on each component > 0.035/< -0.035) up regulated proteins (in Alcoholism, logFC>0) from all three compartments for both components, and performed ORA analysis to look for KEGG pathways that those up-regulated proteins enriched in. 

* Meanwhile, we also combine top (feature loading on each component < -0.035/< -0.035) down regulated proteins (in Alcoholism, logFC<0) from all three compartments for both components, and performed ORA analysis to look for KEGG pathways that those down-regulated proteins enriched in. 

```{r, message=FALSE, warning=FALSE}
# final.models = readRDS(paste0(OUTDIR, "final_models.RDS"))
compartments = c("CYTOSOLIC", "PELLET", "SYNAPTOSOMAL")
names = c("CYTOSOLIC", "NUCLEAR", "SYNAPTOSOMAL")
loading_cutoff = 0.035
library(org.Hs.eg.db)
i = 1
up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.> loading_cutoff] %>% names()
up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.< -loading_cutoff] %>% names()
C = merge(final.models[[i]]$loadings$X[up1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[up2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")

i = 2
up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.< -loading_cutoff] %>% names()
up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.> loading_cutoff] %>% names()
I = merge(final.models[[i]]$loadings$X[up1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[up2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")

i = 3
up1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.< -loading_cutoff] %>% names()
up2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.< -loading_cutoff] %>% names()
S = merge(final.models[[i]]$loadings$X[up1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[up2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")
  
up = list(C,I,S)
up = lapply(seq_along(compartments), function(i) {
  top = up[[i]]
  top[intersect(rownames(top),protein.list$dir.list[[i]][[1]]),,drop=F] %>%
    as.data.frame() %>% mutate("Reg"="up", "Fraction"=names[i])
}) %>% setNames(compartments)
# Combine the DEG result together to get all the protein_IDs and genes
all = Reduce(full_join, rslt.list)
# dim(all)
up.genes = all[all$Protein_IDs%in%(Reduce(rbind, up) %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 1
up.genes_Cytosolic = all[all$Protein_IDs%in%(up[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 2
up.genes_Insoluble = all[all$Protein_IDs%in%(up[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 3
up.genes_Synaptosomal = all[all$Protein_IDs%in%(up[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
  

i = 1
down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.< -loading_cutoff] %>% names()
down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.> loading_cutoff] %>% names()
C = merge(final.models[[i]]$loadings$X[down1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[down2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")

i = 2
down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.> loading_cutoff] %>% names()
down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.< -loading_cutoff] %>% names()
I = merge(final.models[[i]]$loadings$X[down1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[down2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")

i = 3
down1 = final.models[[i]]$loadings$X[, "comp1"] %>%
  .[.> loading_cutoff] %>% names()
down2 = final.models[[i]]$loadings$X[, "comp2"] %>%
  .[.> loading_cutoff] %>% names()
S = merge(final.models[[i]]$loadings$X[down1,"comp1",drop=F] %>% as.data.frame(), 
              final.models[[i]]$loadings$X[down2,"comp2",drop=F] %>% as.data.frame(),
          by=0, all=TRUE) %>% column_to_rownames(var="Row.names")
down = list(C, I, S)
down = lapply(seq_along(compartments), function(i) {
  top = down[[i]]
  top[intersect(rownames(top),protein.list$dir.list[[i]][[2]]),,drop=F] %>%
    as.data.frame() %>% mutate("Reg"="down", "Fraction"=names[i])
}) %>% setNames(compartments)
down.genes = all[all$Protein_IDs%in%(Reduce(rbind, down) %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 1
down.genes_Cytosolic = all[all$Protein_IDs%in%(down[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 2
down.genes_Insoluble = all[all$Protein_IDs%in%(down[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")
i = 3
down.genes_Synaptosomal = all[all$Protein_IDs%in%(down[[i]] %>% 
                rownames() %>% unique()), "Genes", drop=T] %>%
                unique() %>% 
                mapIds(org.Hs.eg.db, keys = ., 
                column = "ENTREZID", keytype = "SYMBOL", 
                multiVals = "first")


entrez_list_sig = list(up = up.genes, down = down.genes)
entrez_list_sig_Cytosolic = list(up = up.genes_Cytosolic, down = down.genes_Cytosolic)
entrez_list_sig_Insoluble = list(up = up.genes_Insoluble, down = down.genes_Insoluble)
entrez_list_sig_Synaptosomal = list(up = up.genes_Synaptosomal, down = down.genes_Synaptosomal)
saveRDS(list(entrez_list_sig=entrez_list_sig,
             entrez_list_sig_Cytosolic=entrez_list_sig_Cytosolic,
             entrez_list_sig_Insoluble=entrez_list_sig_Insoluble,
             entrez_list_sig_Synaptosomal=entrez_list_sig_Synaptosomal),
        paste0(OUTDIR, "/pathwayAnalysis/entrez_list.RDS"))

# Write the top up and top down proteins and their loadings into an excel file
dir.create(paste0(OUTDIR, "/pathwayAnalysis/"), recursive = T, showWarnings = F)
options(java.parameters = "-Xmx1000m")
up = lapply(up, function(u) {
  as.data.frame(u) %>%
    rownames_to_column(var="Protein_IDs")
})
down = lapply(down, function(d) {
  as.data.frame(d) %>%
    rownames_to_column(var="Protein_IDs")
})
write.xlsx(as.data.frame(rbind(Reduce(rbind, up), 
                               Reduce(rbind, down))), 
             paste0(OUTDIR, "/pathwayAnalysis/top_loading_proteins_used4ORA.xlsx"),
             col.names=TRUE, row.names=F,
             append=F)

# Add gene names
# library(readxl)
# OUTDIR = "../results/rerun/plsDA/"
# compartments = c("CYTOSOLIC", "PELLET", "SYNAPTOSOMAL")
# rslt.list = lapply(compartments, function(i) {
#   read_excel(
#     paste0("../results/rerun/DifferentialAnalysis/Differential_Table_",i,".xlsx"), 
#                   sheet=1, col_names=TRUE, na="", skip=0)
# }) %>% setNames(compartments)
# all = Reduce(full_join, rslt.list)
keptLoading = read_excel(
    paste0(OUTDIR, "/pathwayAnalysis/top_loading_proteins_used4ORA.xlsx"), 
                  sheet=1, col_names=TRUE, na="", skip=0)
# colnames(keptLoading)[1] = "Protein_IDs"
mapping = setNames(all[!duplicated(all$Protein_IDs), "Genes", drop=T], 
         all[!duplicated(all$Protein_IDs), "Protein_IDs", drop=T])
keptLoading$Genes = mapping[keptLoading$Protein_IDs] %>% as.character()
# keptLoading = tibble::column_to_rownames(keptLoading, var="Protein_IDs")
write.xlsx(as.data.frame(keptLoading), 
             paste0(OUTDIR, "/pathwayAnalysis/top_loading_proteins_used4ORA.xlsx"),
             col.names=TRUE, row.names=F,
             append=F)

```

```{r, meassage=FALSE, warning=FALSE}
library(clusterProfiler)
# print("KEGG GSEA analysis")
# library(xlsx)

get.kk.sig = function(entrez_list_sig, NAME, universe) {
  if (file.exists(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_", NAME, ".xlsx"))) {
        file.remove(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_", NAME, ".xlsx"))}
  kk.sig = lapply(names(entrez_list_sig), function(i) {
    gl = entrez_list_sig[[i]] %>% as.character()
    kk = NULL
    if (length(gl)>0) {
      set.seed(1234)
      kk= enrichKEGG(gene = gl,
                     organism = 'hsa',
                     pvalueCutoff = 1,
                     qvalueCutoff = 1,
                     universe = universe,
                     use_internal_data = F)
      print("Convert Entrez ID back to gene symbol")
      source("../src/convertEntrez2Symbol.R")
      kk@result$geneID_Symbol = 
        convertEntrez2Symbol(kk@result$geneID, mapping_list=entrez_list_sig[[i]]) %>%
        as.character()
      options(java.parameters = "-Xmx1000m")
      write.xlsx(as.data.frame(kk@result), 
                 paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_", NAME, ".xlsx"),
                 sheetName = paste0(i, "_plsDA"),
                 col.names=TRUE, row.names=FALSE,
                 append=T)
    }
    kk
  }) %>% setNames(names(entrez_list_sig)) 
  return(kk.sig)
}

universe = lapply(normData, function(n){
  n$Genes %>% as.character()
}) %>% unlist() %>% unique() %>% bitr(
  .,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Hs.eg.db) %>%
  pull(ENTREZID) %>% as.character() %>% unique() %>% na.omit()

NAME = "ALL"
kk.sig = get.kk.sig(entrez_list_sig, NAME=NAME, universe=universe)
saveRDS(kk.sig, paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))
# kk.sig = readRDS(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))

NAME = "Cytosolic"
kk.sig_Cytosolic = get.kk.sig(entrez_list_sig_Cytosolic, NAME=NAME, universe=universe)
saveRDS(kk.sig_Cytosolic, paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))
# kk.sig_Cytosolic = readRDS(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))

NAME = "Insoluble"
kk.sig_Insoluble = get.kk.sig(entrez_list_sig_Insoluble, NAME=NAME, universe=universe)
saveRDS(kk.sig_Insoluble, paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))
# kk.sig_Insoluble = readRDS(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))

NAME = "Synaptosomal"
kk.sig_Synaptosomal = get.kk.sig(entrez_list_sig_Synaptosomal, NAME=NAME, universe=universe)
saveRDS(kk.sig_Synaptosomal, paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))
# kk.sig_Synaptosomal = readRDS(paste0(OUTDIR, "pathwayAnalysis/KEGG_ORA_result_", NAME, ".RDS"))

```

### Dotplot showing significant KEGG pathways

```{r, message=FALSE, warning=FALSE}
source("../src/makeDotplot.R")

qval.thresh = c(0.1, 0.05)
lapply(qval.thresh, function(q) {
    pdf(paste0(OUTDIR, "/pathwayAnalysis/KEGG_Pathways_DotPlot_p.adjustCutoff_", q, "_all.pdf"), h=2.9, w=6.5)
      print(makeDotplot_ORA_KEGG(
        list(ALL=kk.sig), 
        index.list=1, 
        p.adjust.thresh=q))
    invisible(dev.off()) 
})

q = 0.1
pdf(paste0(OUTDIR, "/pathwayAnalysis/KEGG_Pathways_DotPlot_p.adjustCutoff_", q, "_separately.pdf"), h=10, w=9)
  print(makeDotplot_ORA_KEGG(
    list(Cytosolic=kk.sig_Cytosolic, Nuclear=kk.sig_Insoluble, Synaptosomal=kk.sig_Synaptosomal), 
    index.list=1:3, 
    p.adjust.thresh=q))
invisible(dev.off()) 

q = 0.05
pdf(paste0(OUTDIR, "/pathwayAnalysis/KEGG_Pathways_DotPlot_p.adjustCutoff_", q, "_separately.pdf"), h=6.5, w=9)
  print(makeDotplot_ORA_KEGG(
    list(Cytosolic=kk.sig_Cytosolic, Nuclear=kk.sig_Insoluble, Synaptosomal=kk.sig_Synaptosomal), 
    index.list=1:3, 
    p.adjust.thresh=q))
invisible(dev.off()) 

```

## ORA on synGO

```{r}
# Import plsDA selected genes
plsDA_out_syn = 
  readxl::read_excel("../results/rerun/plsDA/pathwayAnalysis/top_loading_proteins_used4ORA.xlsx") %>% 
  filter(Fraction=="SYNAPTOSOMAL")
up = filter(plsDA_out_syn, Reg=="up") %>% pull(Genes)
down = filter(plsDA_out_syn, Reg=="down") %>% pull(Genes)
all = c(up, down)
up
down
length(all)

# synGO
synGO = xlsx::read.xlsx2("../data/SynGO_bulk_download_release_20231201/syngo_ontologies.xlsx",
                   sheetName = "Sheet 1", header=T) #%>% filter(domain=="BP")
syn_ref = split(synGO$hgnc_symbol, synGO$name) 
syn_set = lapply(names(syn_ref), function(s) {
    set = syn_ref[[s]]
    as.data.frame(list(gs_name = s, human_entrez_gene=strsplit(set, ", ")[[1]] %>% unique()))
  }) %>% Reduce(rbind, .) 


normData = 
  readRDS("../results/rerun/QC/NormalizedIntensity_dataTable.list_SL_IRS_TMM.RDS")

# Run ORA on synGO
library(clusterProfiler)
synGO_ora_results_up = enricher(
  gene = up, # A vector of your genes of interest
  pvalueCutoff = 0.1, # Can choose a FDR cutoff
  pAdjustMethod = "BH", # Method to be used for multiple testing correction
  universe = normData$SYNAPTOSOMAL$Genes%>%unique(), # A vector containing your background set genes
  # The pathway information should be a data frame with a term name or
  # identifier and the gene identifiers
  TERM2GENE = syn_set%>%as.data.frame()
  )
synGO_ora_results_up@result

synGO_ora_results_down = enricher(
  gene = up, # A vector of your genes of interest
  pvalueCutoff = 0.1, # Can choose a FDR cutoff
  pAdjustMethod = "BH", # Method to be used for multiple testing correction
  universe = normData$SYNAPTOSOMAL$Genes%>%unique(), # A vector containing your background set genes
  # The pathway information should be a data frame with a term name or
  # identifier and the gene identifiers
  TERM2GENE = syn_set%>%as.data.frame()
  )
synGO_ora_results_down@result

TableNames = list("Synaptosomal_up"=synGO_ora_results_up@result,
                  "Synaptosomal_down"=synGO_ora_results_down@result)
openxlsx::write.xlsx(TableNames, "../results/rerun/plsDA/pathwayAnalysis/synGO/")


```



# Spliceosome protein visualization

```{r, fig.height=8, fig.width=8}
# Proteins/genes
Spliceosome_genes = kk.sig_Insoluble$up@result %>% 
  filter(Description=="Spliceosome") %>% 
  pull(geneID_Symbol) %>% str_split("/") %>% 
  .[[1]]

# Design
design = read_table("../data/design_7314/design_PELLET.tsv")
design

# Normalized protein abundance
Insoluble_Nintensity = read_excel(
    paste0("../results/rerun/QC/NormalizedIntensity.xlsx"), 
                  sheet=2, col_names=TRUE, na="", skip=0) %>%
    filter(Genes %in% Spliceosome_genes) %>%
    gather("sample", "Norm.intensity", -Genes, -Gene_names, -Protein_IDs) %>%
    full_join(design)
Insoluble_Nintensity$Condition = ifelse(Insoluble_Nintensity$Condition=="case", "AUD", "UC")

theme_set(theme_bw())
P = ggplot(Insoluble_Nintensity, aes(x=Condition, y=Norm.intensity)) +
  geom_violin(aes(fill=Condition)) + 
  facet_wrap(~Protein_IDs, nrow=4, scale="free") +
  geom_boxplot(width=0.1) +
  geom_point(position = position_jitter(seed = 1, width = 0.2))
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=1)

P

pdf(paste0(OUTDIR, "/pathwayAnalysis/Spliceosome_proteins_visualization.pdf"), h=6.5, w=8)
P
dev.off()
```



# Session Info

```{r}
out = capture.output(print(sessionInfo()))
writeLines(out, paste0(OUTDIR, "sessionInfo.txt"))

print(sessionInfo())

```

# Deliverables

# Directory

## Practical info

## Software

# Data responsibility

The responsibility for data archiving lies wscale = TRUEith the PI of the project. We do not offer long-term storage or retrieval of data.

-   **NBIS & UPPNEX:** We kindly ask that you remove the files from UPPMAX/UPPNEX. The main storage at UPPNEX is optimized for high-speed and parallel access, which makes it expensive and not the right place for longer time archiving. Please consider others by not taking up the expensive space. Please note that UPPMAX is a resource separate from the Bioinformatics Platform, administered by the Swedish National Infrastructure for Computing (SNIC) and SNIC-specifc project rules apply to all projects hosted at UPPMAX.
-   **Sensitive data:** Please note that special considerations may apply to the human-derived legally considered sensitive personal data. These should be handled according to specific laws and regulations as outlined e.g. [here](http://nbis.se/support/human-data.html).
-   **Long-term backup:** We recommend asking your local IT for support with long-term data archiving. The [Data Office](https://www.scilifelab.se/data/) at SciLifeLab may be of help to discuss other options.

# Acknowledgments

If you are presenting the results in a paper, at a workshop or conference, we kindly ask you to acknowledge us.

-   NBIS staff are encouraged to be co-authors when this is merited in accordance to the ethical recommendations for authorship, e.g. [ICMJE recommendations](http://www.icmje.org/recommendations/browse/roles-and-responsibilities/defining-the-role-of-authors-and-contributors.html). If applicable, please include **Name, Surname, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Further Affliations**, as co-author. In other cases, NBIS would be grateful if support by us is acknowledged in publications according to this example:

> "Support by NBIS (National Bioinformatics Infrastructure Sweden) is gratefully acknowledged."

-   **UPPMAX** kindly asks you to [acknowledge UPPMAX and NAISS](https://www.uppmax.uu.se/support/faq/general-miscellaneous-faq/acknowledging-uppmax-and-naiss).

-   **NGI:** For publications based on data from NGI Sweden, NGI and SciLifeLab should be acknowledged. Instructions are included in the NGI reports as described [here](https://ngisweden.scilifelab.se/resources/getting-started-at-ngi/).
